Backbone-Store: An API for Automating Local Caching of Web Application Data

### 2012-05-21 paper/paper.tex
final

\begin{abstract}
Web applications that require re-fetching data from a server
can see an improved load time by storing old data in the client's browser,
requiring only new data to be retrieved from the server. In this paper, we present Backbone-Store: a local caching API that extends Backbone.js and uses the HTML5 localStorage API to offer web developers two different persistence strategies that easily store and reload a web application's data. To evaluate the API,  we analyze three simple applications to measure the performance characteristics of loading documents, medium-sized emails, and small notes from both the local storage on the client's browser and from the server. As expected, we find that our two implementations of local storage both perform faster than loading the same data from the server, with the one-key-per-collection persistence strategy outperforming the one-key-per-model strategy. This time difference increases for larger quantities of the same type of item, suggesting that local storage becomes more advantageous as the amount of data loaded by an application increases. Furthermore, we examine the time required for an application to save data to the local storage, comparing the relative save-time performance of our two persistence strategies and finding one to be more suitable for small updates to the application's data and the other to be more suitable for larger updates.
\end{abstract}

@@

HTML5 presents new options for storing a web application's data in the client's
browser. Efficient local storage presents two major
benefits. The first benefit is that web applications can store information that is loaded and saved
frequently to help speed up the time it takes to present the user
interface. Users are sensitive to the time it takes a
web application to load, with slower applications causing users to navigate
away or think less of the site. The second benefit is that the web applications
can request less data from a server if it can store information that it
previously would have needed to fetch from the server. This allows a server to
fulfill more requests and thus reduces costs to maintain servers to meet a web
application's requests.

@@

Backbone-store, the project described in this paper, presents a new way for web
application programmers to easily save their web application's data on the
client side to take advantage of the benefits of local storage (as described in the Introduction). This project is an extension of the pure-Javascript Backbone.js \cite{backbone} library, which is used by many companies (e.g., Four Square, Walmart Mobile, Pandora, Trello, etc.) to create web applications for browsers and mobile devices. Data
stored in Backbone.js applications are generally placed in either a Model
object or a Collection object, which is an ordered list of objects of one type
of Model. Backbone.js also provides View objects for managing the UI based on
the state of the Model or Collection objects. We used Backbone.js as a base due
to its aforementioned popularity with many web developers, and because it
provides a very good model for developers that is easy to modify with our extensions.

Currently, web programmers wishing to employ local caching using Backbone.js would need to write a custom storage solution (including custom libraries and scripts tailored to the particular application) to save the data from each of their models and/or collections to HTML5's localStorage or another local storage option. Backbone-store simplifies this process by providing an API that automates the local caching process such that information from a saved model or collection can be reloaded into the
models and collections the next time the web application starts. With a few new
methods added to the Model and Collection objects, a web developer can choose
when a copy of a model's or collection's data should be stored, and when to
load saved data back into the application's models or collections. There is
also a choice of how the data is stored for collections of models. All models
can be stored as a bundle under one key in the localStorage (Strategy 1), or they can be
stored individually, with each model under its own key in the localStorage (Strategy 2).
These two persistence strategies provide flexibility for different use cases where one option may be
significantly better than the other, depending on the number of items and
the amount of data being stored. We evaluate the performance of these two strategies in the Results section.

@@

The first main problem is the diversity of data types stored by web applications. Indeed, Web applications vary in both the quantity and size of data item they must request from a
server before loading the user interface, causing possible variance in the load and save times of various persistence strategies.  For example, some applications might require only a few
large items, such as showing a user a menu of large documents they can edit. On
the opposite end of both the quantity and size spectra, a note-taking application might request many
relatively small notes. We address the varying data characteristics by evaluating the load and save times of each persistent strategy for applications using data of varying quantities and sizes and highlight cases where each strategy prevails.

@@

Our project makes two main contributions. First, we create an API that
automates local caching for web applications. This API will allow developers to
choose from two persistence strategies for locally cached data:

@@

Second, we conduct a performance evaluation of load times for data
stored locally (using the aforementioned two strategies) compared to data
stored remotely for three types of web applications that differ in item size
per item (70 bytes, 900 bytes, 3200 bytes) and item quantity (10, 50, 100, 200,
300 items per collection) and highlight conditions where local storage is most
advantageous.

@@

Note that Step 1 and 2 are separated by a delay so as not to overlap in
execution and skew the load time.

@@

It is of note that users typically perceive lag in an interface when an
operation takes more than 100 milliseconds to complete. In all of our tests,
the one-key-per-collection persistence strategy was able to keep the load time for the
collection below or very near 100 ms, meaning that users would perceive the
load time as nearly instantaneous. On the other hand, the load time from the server is much
higher, so users would perceive noticeable lag while waiting for the application
to load. This demonstrates that our extension provides application users with
an improved user experience, due to noticeably faster application load times. Additionally, moving storage from server to client will also reduce the load on the server, allowing it to take more requests from applications, further enhancing its performance and speed.

@@

This report describes Backbone-Store: our extension for the Backbone.js library, which
simplifies the task of persisting a collection of data models to a web
browser's local storage, in order to facilitate faster load times and reduce
server load. This extension provides an API that automatically saves and loads collections on demand, using one of two different persistence strategies: one-key-per-model and one-key-per-collection. We test our extension using three different demo applications (Documents, Emails, Notes), with different characteristics for the data that are being used by the application. Our tests
vary both the number of data items in a single collection and the size of
each data item, to simulate a number of different application scenarios. Finally, our
test applications show that Backbone-Store is effectively and significantly reduces the load times in many different scenarios, resulting in an improved experience for the application
user and a more convenient, simpler programming experience for application developers.

>>>>

\begin{abstract}
Web applications that require multiple fetches of the data from the server
can see an improved load time by storing old data in the client's browser,
requiring only new data to be retrieved from the server. In this paper, we present Backbone-Store: a local caching API that extends Backbone.js and uses the HTML5 localStorage API to provide web developers with two different persistence strategies to easily store and reload an application's data. To evaluate the API,  we analyze three simple test applications to measure the performance characteristics of loading documents, medium-sized emails, and small notes from both the local storage on the client's browser and from the server. As expected, we find that both persistence strategies perform faster than loading the same data from the server, with the one-key-per-collection strategy outperforming the one-key-per-model strategy. This time difference increases for larger quantities of the same type of item, suggesting that local storage becomes more advantageous as the number of items loaded by an application increases. Furthermore, we compare the relative saving time performance of our two persistence strategies and find one to be more suitable for small updates to the application's data and the other to be more suitable for larger updates.
\end{abstract}

@@

HTML5 presents new options for storing a web application's data in the client's
browser. Efficient local storage presents two major
benefits. The first benefit is that web applications can store information that is loaded and saved
frequently to help speed up the time it takes to displaythe user
interface. Users are sensitive to the time it takes a
web application to load, with slower applications causing users to navigate
away or think less of the site. The second benefit is that the web applications
can request less data from a server if it can store information that it
previously would have needed to fetch from the server. This allows a server to
fulfill more requests and thus reduces costs to maintain servers to meet a web
application's requests.

@@

Backbone-store, the project described in this paper, presents a new way for web
application programmers to easily save their web application's data on the
client side to take advantage of the benefits of local storage (as described in the Introduction). This project is an extension of the pure-Javascript Backbone.js \cite{backbone} library, which is used by many companies (e.g., Four Square, Walmart Mobile, Pandora, Trello, etc.) to create web applications for browsers and mobile devices. Data
stored in Backbone.js applications are generally placed in either a Model
object or a Collection object, which is an ordered list of objects of one type
of Model. Backbone.js also provides View objects for managing the UI based on
the state of the Model or Collection objects. We use Backbone.js as a base due
to its aforementioned popularity with many web developers, and because it
provides a very good model for developers that is easy to modify with our extensions.

Currently, web programmers wishing to employ local caching using Backbone.js need to write a custom storage solution (including custom libraries and scripts tailored to the particular application) to save the data from each of their models and/or collections to HTML5's localStorage or another local storage option. Backbone-store simplifies this process by providing an API that automates the local caching process such that information from a saved model or collection can be reloaded into the
models and collections the next time the web application starts. With a few new
methods added to the Model and Collection objects, a web developer can choose
when a copy of a model's or collection's data should be stored, and when to
load saved data back into the application's models or collections. There is
also a choice of how the data is stored for collections of models. All models
can be stored as a bundle under one key in the localStorage (Strategy 1), or they can be
stored individually, with each model under its own key in the localStorage (Strategy 2).
These two persistence strategies provide flexibility for different use cases where one option may be
significantly better than the other, depending on the number of items and
the amount of data being stored. We evaluate the performance of these two strategies in the Results section.

@@

The first main problem is the diversity of data stored by web applications. Indeed, Web applications vary in both the quantity and size of data item they must request from a
server before loading the user interface, causing possible variance in the load and save times of various persistence strategies.  For example, some applications might require only a few
large items, such as showing a user a menu of large documents they can edit. On
opposite ends of the quantity and size spectra, a note-taking application might request many
relatively small notes. We address the varying data characteristics by evaluating the load and save times of each persistent strategy for applications using data of varying quantities and sizes and highlight cases where each strategy prevails.

@@

To prove our hypothesis, this project makes two main contributions. First, we create an API that
automates local caching for web applications. This API will allow developers to
choose from two persistence strategies for locally cached data:

@@

Second, we conduct a performance evaluation of load times for data
stored locally (using the aforementioned two strategies) compared to data
stored remotely for three types of web applications that differ in item size
per item (70 bytes, 900 bytes, 3200 bytes) and item quantity (10, 50, 100, 200,
300 items per collection) and highlight conditions where local storage is most
advantageous. We also evaluate how long it takes for a collection of varying numbers of items to be saved to local storage, noting which persistence strategy performs better.

@@

(Note that Step 1 and 2 are separated by a delay so as not to overlap in
execution and skew the load time.)

@@

It is of note that users typically perceive lag in an interface when an
operation takes more than 100 milliseconds to complete. In all of our tests,
the one-key-per-collection persistence strategy was able to keep the load time for the
collection below or very near 100 ms, meaning that users would perceive the
load time as nearly instantaneous. On the other hand, the load time from the server is much
higher, so users would perceive noticeable lag while waiting for the application
to load. This demonstrates that our extension provides application users with
an improved user experience, due to noticeably faster application load times. Additionally, moving storage from server to client will reduce the load on the server, allowing it to take more requests from applications, further enhancing its performance and speed.

@@

This report describes Backbone-Store: our extension for the Backbone.js library, which
simplifies the task of persisting a collection of data models to a web
browser's local storage, in order to facilitate faster load times and reduce
server load. This extension provides an API that automatically saves and loads collections on demand, using one of two different persistence strategies: one-key-per-model and one-key-per-collection. We test our extension using three different demo applications (Documents, Emails, Notes), with different characteristics for the data that are being used by the application. Our tests
vary both the number of data items in a single collection and the size of
each data item, to simulate a number of different application scenarios. Finally, our
test applications show that Backbone-Store effectively and significantly reduces the load times in many different scenarios, resulting in an improved experience for the application
user while simplifying the local caching programming process for application developers.
